---
title: "Mood"
author: "Arda Ergin"
date: "2024-04-02"
output: html_document
---

# Setup
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(scales)
library(lubridate)
```

```{r}
data_raw <- read.csv("data.csv")
```

# Start

How many variables and how many participants?
```{r}
unique(data_raw$variable)
unique(data_raw$id)
```

Some Fixes:
```{r}
# Getting a better ID column:
data_0 <- data_raw %>%
  mutate(ID = as.numeric(sapply(
    strsplit(id, split = '\\.'), function(x) x[2])))

# Changing the name for the "X" variable
colnames(data_0)[1] <- "obs_number"

# Getting rid of the 
data_1 <- data_0[,c("obs_number", "time", "ID", "variable", "value")]
```

Long to Wide Transformation:
```{r}
wide_data <- tidyr::pivot_wider(
  data_0, 
  names_from = variable, 
  values_from = value)

data <- as.data.frame(wide_data)
```

Looking at Just one Individual
```{r}
data_id_1 <- data %>% filter(id == "AS14.01")
```







# Investigating Variables
## time
```{r}
# Checking if there is any NAs in the `time` variable
any(is.na(data$time))

# Transforming the variable into an time variable:
data <- data %>% mutate(
  date_time = as.POSIXct(
    as.character(time), 
    format="%Y-%m-%d %H:%M:%OS"))
data <- subset(data, select = -time)

# Quick Summary
summary(data$date_time)
```


### New Variables
```{r}
# Creating Variable: the Date (of Observation)
data <- data %>%
  mutate(date = as.Date(date_time))

# Creating Variable: the Time (of Observation)
data <- data %>%
  mutate(time = format(date_time, "%H:%M:%S"))

# Creating Variable: the Hour (of Observation)
data <- data %>%
  mutate(hour = format(date_time, "%H"))

# Creating Variable: the day of the week (of Observation)
data <- data %>%
  mutate(weekday = weekdays(date_time),
         weekday = factor(
           weekday, 
           levels = c("Monday", "Tuesday", 
                      "Wednesday", "Thursday", 
                      "Friday", "Saturday", 
                      "Sunday")))

# Creating Variable: the week of the year (of Observation)
data <- data %>%
  mutate(week_of_year = lubridate::week(date_time))
```


### Visual: Entire Time
```{r}
# Histogram
data %>% ggplot(aes(x = time_t)) + 
  geom_histogram(bins = 100, 
                 color = "black", 
                 fill = "purple2") +
  theme_classic(base_family = "Times") + 
  labs(title = "Distribution of Observation Times", 
       x = "Date-Time", 
       y = "Frequency") +
  # Rotating x-axis labels for better readability:
  theme(axis.text.x = element_text(
    angle = 30, hjust = 1)) +
  scale_x_datetime(
    date_breaks = "2 weeks", 
    date_labels = "%d %b %Y")
```

### Visual: Weekdays
```{r}
# Summarizing Per Weekday
weekday_summary <- data %>%
  group_by(weekday) %>%
  summarise(
    mean_mood = mean(mood, na.rm = TRUE),
    sd_mood = sd(mood, na.rm = TRUE),
    n = sum(!is.na(mood)),
    sem = sd_mood / sqrt(n),
    ci_lower = mean_mood - qnorm(0.975) * sem,
    ci_upper = mean_mood + qnorm(0.975) * sem
  )

# Bar Plot
weekday_summary %>%
  ggplot(aes(
    x = weekday, 
    y = mean_mood)) + 
  geom_col(color = "black", 
           fill = "skyblue") +
  geom_errorbar(aes(ymin = ci_lower, 
                    ymax = ci_upper), 
                width = .2) +  
  geom_text(aes(
    label = n, 
    y = mean_mood + 0.02), 
    vjust = -5.5,
    family = "Times New Roman") + 
  theme_classic(base_family = "Times") +
  labs(
    title = "Mean Mood per Weekday",
    x = "Weekday",
    y = "Mean Mood"
  ) + 
  coord_cartesian(ylim = c(6.8,7.5))
  
```


### Visual: Week of Year
```{r}
# Summarizing Per Weekday
week_summary <- data %>%
  group_by(week_of_year) %>%
  summarise(
    mean_mood = mean(mood, na.rm = TRUE),
    sd_mood = sd(mood, na.rm = TRUE),
    n = sum(!is.na(mood)),
    sem = sd_mood / sqrt(n),
    ci_lower = mean_mood - qnorm(0.975) * sem,
    ci_upper = mean_mood + qnorm(0.975) * sem
  )

# Bar Plot
week_summary %>%
  ggplot(aes(
    x = week_of_year, 
    y = mean_mood)) + 
  geom_col(color = "black", 
           fill = "skyblue") +
  geom_errorbar(aes(ymin = ci_lower, 
                    ymax = ci_upper), 
                width = .2) +  
  geom_text(aes(
    label = n, 
    y = mean_mood + 0.02), 
    vjust = -5.5,
    family = "Times New Roman") + 
  theme_classic(base_family = "Times") +
  labs(
    title = "Mean Mood per Week of the Year",
    x = "Week of the Year",
    y = "Mean Mood"
  ) + 
  coord_cartesian(ylim = c(6,7.6))
  
```


```{r}
x <- data %>%
  mutate(
    time_t = as.POSIXct(as.character(time), format="%Y-%m-%d %H:%M:%OS"),
    time_only = format(time_t, "%H:%M:%S"),
    seconds_since_midnight = hour(time_t) * 3600 + minute(time_t) * 60 + second(time_t)
  )

# Now you can create a histogram of `seconds_since_midnight`
hist(x$seconds_since_midnight, breaks=24, main="Distribution of Time Throughout the Day",
     xlab="Seconds Since Midnight", ylab="Frequency")

```



## Mood
```{r}
# Quick Summary
summary(wide_data$mood)
table(wide_data$mood)

# Histogram
data %>% ggplot(aes(x = mood)) + 
  geom_histogram(bins = 10, color = "black", fill = "firebrick") +
  theme_classic(base_family = "Times") + 
  labs(title = "Distribution of Mood Observations", 
       x = "Mood", 
       y = "Frequency")
```

```{r}
mood_summary <- data %>%
  group_by(ID) %>%
  summarise(mean_mood = mean(mood, na.rm = TRUE),
            median_mood = median(mood, na.rm = TRUE),
            sd_mood = sd(mood, na.rm = TRUE),
            max_mood = max(mood, na.rm = TRUE),
            min_mood = min(mood, na.rm = TRUE))
mood_summary
```

Important the understand our DV, so good to summarize is for each participant.        
- Overall, there does not seem to be much of a difference between the individuals in terms of their mean and median.
- On the other hand, mood fluctuations are not the same for all individuals, and there might be individuals who experience more fluctuations.

```{r}
for (participant in unique(data$ID)) {
  moods <- data %>%
    filter(!is.na(mood))
  
  mood <- moods %>% filter(ID == participant)
  
  p <- mood %>% ggplot(
    aes(x = time_t, 
        y = mood)) +
    geom_line(color = colors()[sample(8:length(colors()),1)]) + 
    scale_y_continuous(limits = c(1, 10)) +
    theme_classic(base_family = "Times") + 
    theme(legend.position = "none") +
    labs(title = paste(
      "Mood Over Time for Participant Number", 
      participant), 
         x = "Time", 
         y = "Mood")
  
  print(p)
}
```

Participant number 7 have too much fluctuations and can be considered an outlier among the other participants. It might possibly be a good idea to exclude them. Maybe also participant number 33.

```{r}
# Creating a new variable in our "mood_summary" 
# Putting the ID to the column if the participant is an outlier 
# FROM: https://www.r-bloggers.com/2022/08/how-to-label-outliers-in-boxplots-in-ggplot2/
find_outlier <- function(x) {
  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
}

mood_summary <- mood_summary %>%
  mutate(outlier = ifelse(find_outlier(sd_mood), ID, NA))

mood_summary %>% ggplot(aes(x = "", y = sd_mood)) +
  geom_boxplot(color = "firebrick") +
  theme_classic() +
  labs(title = "Boxplot of Standard Deviation of Mood",
       y = "Standard Deviation of Mood",
       x = "") + 
  geom_text(aes(label=outlier), na.rm=TRUE, hjust=-1)
```

Although, it needs to be noted that individuals who have depression will show these mood fluctuations. If we assume that this sample is a representative sample, considering the $10-20\%$ prevalence, we should expect around $2-4$ individuals that are displaying such patterns. Hence, it is a question whether this should actually be considered as an "anomaly". 





## Arousal & Valence
```{r}
# Quick Summary: Arousal
summary(data$circumplex.arousal)
table(data$circumplex.arousal)

# Quick Summary: Valence
summary(data$circumplex.valence)
table(data$circumplex.valence)

# Histograms & Combining Them
hist_arousal <- data %>% ggplot(aes(x = circumplex.arousal)) + 
  geom_histogram(bins = 5, color = "black", fill = "deepskyblue2") +
  theme_classic(base_family = "Times") + 
  labs(title = "Distribution of Arousal", 
       x = "Arousal", 
       y = "Frequency")

hist_valence <- data %>% ggplot(aes(x = circumplex.valence)) + 
  geom_histogram(bins = 5, color = "black", fill = "purple2") +
  theme_classic(base_family = "Times") + 
  labs(title = "Distribution of Valence", 
       x = "Valence", 
       y = "Frequency")

# Doing layout for ggpplot2:
(hist_arousal | hist_valence) + 
  patchwork::plot_layout(ncol = 2)
```


## circumplex.arousal
```{r}
hist(wide_data$circumplex.arousal)
summary(wide_data$circumplex.arousal)
table(wide_data$circumplex.arousal)
```

## circumplex.valence
```{r}
hist(wide_data$circumplex.valence)
summary(wide_data$circumplex.valence)
table(wide_data$circumplex.valence)

names(wide_data)
```


```{r}
hist(log(wide_data$activity))
summary(wide_data$activity)
qqnorm(log(wide_data$activity))
```


## appCat.game
```{r}
gamers <- wide_data %>% filter(appCat.game > 300)
hist(log(wide_data$appCat.game), breaks = 100)
boxplot(log(wide_data$appCat.game))
```


## appCat.game

```{r}
max(wide_data$appCat.game)
summary(wide_data$sms)
```


# Data Analysis

```{r}
data_1 <- data %>% filter(ID == 1)

new_data <- data %>%
  group_by(date) %>%
  summarise(
    variable_mean = mean(mood, na.rm = TRUE),
    variable_sum = sum(appCat.game, na.rm = TRUE),
    variable_sum = sum(appCat.social, na.rm = TRUE)
  )
data$appCat.social

x <- lm(variable_mean ~ variable_sum, data = new_data)
summary(x)
```








